---
title: 'Drug Sensitivity Prediction: CCLE Data Set'
author: "Amir Asiaee"
date: "22 January 2020"
output:
  html_document:
    toc: true
    theme: yeti
    highlight: kate
    number_sections: true
---
  
```{r setup, include=FALSE, results="hide"}
knitr::opts_chunk$set(echo = TRUE)
options(width=96)
```
```{r mycss, results="asis", echo=FALSE}
cat('
<style type="text/css">
b, strong {color: red; }
i, em {color: blue; }
.defn {color: purple; }
.para {color: purple;
      font-weight: bold;
}
.figure { text-align: center; }
.caption { font-weight: bold; }
</style>
')
```

# Executive Summary
## Background

## Methods
### Input Data

### Procedure 

### Output Data 

### Statistics

## Results

## Conclusion

# Basics
First we prepare the paths for the clean data. 
```{r pats}
rm(list=ls())
source("00-paths.R")
```


# Loading files
```{r loadData}
load(file.path(paths$clean,"doseResponse.Rda"))
load(file = file.path(paths$clean, "X.Rda"))
```

# Exploring Data
Let's look at the histogram of the cancer types among treated celllines to find the most frequent samples. First we want to check if all of the cell lines have the cancer type include. 

```{r noCancerType}
index <- sapply(strsplit(doseResponse$clName, "_"), function(x) if(length(x) == 1) TRUE else FALSE)
clWithoutType <- unique(doseResponse$clName[index])
```
So there are three cell lines with no cancer type. We remove them. 
```{r removeCL}
dim(doseResponse)
doseResponse <- doseResponse[!(doseResponse$clName %in% clWithoutType),]
dim(doseResponse)
sum(clWithoutType %in% rownames(X))
# rm(index, clWithoutType)
```
It seems that those cellines had no expression measurement and therefore are already removed from X. 

Now, we plot the averge number of each cancer type per drug to see which one of them is useable in our analysis. It seems going below 20 samples is a stretch. So the only useful cancer types are: 
```{r freq}
counts <- table(sapply(strsplit(doseResponse$clName, "_"), function(x) paste(x[2:length(x)], collapse = "_")))
barplot(sort(decreasing = T,  counts / 24), las = 2, cex.names = .75)
print(names(counts)[counts/ 24 > 20])
```


# Generating (X, y, t)
Note that we can do this and all of the following analysis in a single loop. But since we may reuse the data multiple times for various experiment, we are going to do thise separately and save the results. X: measurements, y: response, t: cancer type. 
```{r generateXy}
for(drug in unique(doseResponse$compound)){
  print(paste("Processing", drug))
  y <- doseResponse[doseResponse$compound == drug & doseResponse$clName %in% rownames(X), ]
  xyt <- as.data.frame(X[y$clName, ])
  xyt$yic50 <- y$ic50
  xyt$yaa <- y$actArea
  # extractCancerType
  xyt$t <- sapply(strsplit(y$clName, "_"), function(x) paste(x[2:length(x)], collapse = "_"))
  save(xyt, file = file.path(paths$clean, paste("xyt_", drug, ".Rda", sep = "")))
}
```

Note that because of the way that we made X (from cell line tested at least against one drug) `doseResponse$clName` samples are for sure included in X. But then removed those with zero expression measurements. Therefore, we need the intersection. Also, note that I'm keeping cancer type as character not factor.




































# Appendix: Session Info
This anlysis was performed in this environment:
```{r si}
sessionInfo()
```